---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# leri: R package for the NOAA Landscape Evaporative Response Index

[![CircleCI](https://circleci.com/gh/earthlab/leri/tree/master.svg?style=svg)](https://circleci.com/gh/earthlab/leri/tree/master)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/earthlab/leri?branch=master&svg=true)](https://ci.appveyor.com/project/earthlab/leri)
[![Codecov](https://img.shields.io/codecov/c/github/earthlab/leri.svg)](https://codecov.io/gh/earthlab/leri)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![DOI](https://zenodo.org/badge/160896497.svg)](https://zenodo.org/badge/latestdoi/160896497)


The leri R package facilitates access to the NOAA [Landscape Evaporative 
Response Index](https://www.esrl.noaa.gov/psd/leri/) data product.

## Installation

You can install the development version of leri with devtools:

``` r
# install.packages("devtools")
devtools::install_github("earthlab/leri")
```

Or, you can install the stable version from CRAN:

``` r
install.packages("leri")
```


## Example

The LERI product is available from the year 2000 to present at a 1 km spatial 
resolution over the continental United States, at the following timescales: 

- 1, 3, 7, and 12 month
- 8 day accumulated or non-accumulated during the growing season (April - Oct.)

If we were interested in medium term drought, we could get 3 month timescale
LERI data for a particular date as follows: 

```{r example}
library(leri)
r <- get_leri(date = "2018-09-01", product = "3 month")
r
```

The `get_leri` function will download the NetCDF files from the NOAA ftp server 
and read them as `RasterStack` objects, which facilitates any other operations
that users may need via the `raster` package. 
For example, we could plot the data using the `raster::plot` function: 

```{r plot-leri}
library(raster)
library(viridis)

plot(r, col = cividis(10, direction = -1), 
     main = "3 month LERI data for 2018-09-01")
```

## LERI Resources

More information on the creation and recommended usage for LERI can be found 
here: https://www.esrl.noaa.gov/psd/leri/

Other potentially useful resources include:

- [Poster about the LERI product (pdf)](https://www.esrl.noaa.gov/psd/leri/resources/pdf/LERI_Poster_CPASW-2018_Final.pdf)
- [Slides about the LERI product (pdf)](https://www.esrl.noaa.gov/psd/leri/resources/LERI_Rangwala_NIDIS-BrownBag_092418_PDF.pdf)
